Towards Practical First-Order Model Counting

### 2025-02-06 doc/paper/paper.tex
(Possibly) final major improvements to the paper itself

We end this section by proving that \CompileWithBaseCases terminates since each
recursive call on \autoref{line:final} reduces the number of domains in the
sentence.

\begin{theorem}
  Given any \FO{} sentence $\phi$, \CompileWithBaseCases{$\phi$} terminates.
\end{theorem}

\todo[inline]{Reformulate the theorem above into an upper bound (as I informally
  mention earlier in the text)}

@@

\begin{proof}
  We proceed by induction on the number of domains that variables in $\phi$ are
  quantified over. If there are no domains, then $\phi$ is essentially a
  propositional formula, and \textsc{Crane} compiles it into an equation of the
  form $f = \expr$ with no `function calls'. Suppose that \CompileWithBaseCases
  terminates for all sentences with at most $n \in \mathbb{N}_{0}$ domains. Let
  $\phi$ be a sentence with $n+1$ domains. By \cref{observation1}, each base
  case on \autoref{line:loopy} of \cref{alg:compilewithbasecases} has at least
  one constant. Therefore, by \cref{observation2}, after
  \autoref{line:propagate}, sentence $\psi$ has at most $n$ domains. Thus,
  \autoref{line:final} terminates by the inductive hypothesis, completing the
  proof that \CompileWithBaseCases terminates for an arbitrary sentence with
  $n+1$ domains.
\end{proof}

@@

Future directions for research include conducting a comprehensive experimental
comparison of FOMC algorithms to better understand their comparative performance
across various sentences. The capabilities of \Cranetwo{} could also be
characterised theoretically, e.g., by proving completeness for logic fragments
liftable by other algorithms. Additionally, the efficiency of FOMC algorithms
can be further analysed using fine-grained complexity, which would provide more
detailed insights into the computational demands of different sentences.

\todo[inline]{Cite a bunch of papers for the liftable fragments.}

>>>>

We end this section by bounding the recursion depth of the \CompileWithBaseCases
procedure, thus proving that the algorithm terminates.

\begin{theorem}
  Let $\phi$ be an \FO{} sentence with $n$ domains. Then the maximum recursion
  depth of \CompileWithBaseCases{$\phi$} is $n$.
\end{theorem}

@@

\begin{proof}
  We proceed by induction on $n$. If $n=0$, then $\phi$ is essentially a
  propositional formula, and \textsc{Crane} compiles it into an equation of the
  form $f = \expr$ with no `function calls'. Suppose that---for all sentences
  with at most $n$ domains---\CompileWithBaseCases terminates with a recursion
  depth of at most $n$. Let $\phi$ be a sentence with $n+1$ domains. By
  \cref{observation1}, each base case on \autoref{line:loopy} of
  \cref{alg:compilewithbasecases} has at least one constant. Therefore, by
  \cref{observation2}, after \autoref{line:propagate}, sentence $\psi$ has at
  most $n$ domains. Thus, \autoref{line:final} terminates with a recursion depth
  of at most $n$ by the inductive hypothesis, completing the inductive proof.
\end{proof}

@@

Future directions for research include conducting a comprehensive experimental
comparison of FOMC algorithms to better understand their comparative performance
across various sentences. The capabilities of \Cranetwo{} could also be
characterised theoretically, e.g., by proving completeness for logic fragments
liftable by other
algorithms~\cite{DBLP:journals/jair/Kuzelka21,DBLP:conf/aaai/TothK23,DBLP:journals/ai/BremenK23}.
Additionally, the efficiency of FOMC algorithms can be further analysed using
fine-grained complexity, which would provide more detailed insights into the
computational demands of different sentences.
